---
title: "textables"
author: Bradley Setzler
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This package allows the construction of highly-customized LaTeX table in R. The key functions for creating table rows are:

- `row_numeric`: many options to control numerical formatting. 
    - control decimal places with`dec` argument (e.g., `dec=3` for 3 decimal places);
    - add parentheses (like standard errors) with `se=T`;
    - add percentage symbols with `percentage=T`;
    - add significance stars based on p-values with `pvalues` (e.g., `pvalues=c(0.005,0.05)` would add 3 stars and 2 stars, respectively).
- `row_string`: simple row of strings.
- `row_header`: row that spans all columns except the first.

It also has convenience functions for inserting rules:

- `rule_top`: add a top-rule;
- `rule_mid`: add a mid-rule;
- `rule_bottom`: add a bottom-rule;
- `rule_mid_partial`: add a partial mid-rule;

Finally, it can put the rows together into a tabular or a document with these:

- `tabular_topbottom`: basic tabular environment;
- `tabular_topbottom_pretty`: pretty tabular environment using rules;
- `write_tex`: save as a tex file. The `stand_alone=T` option makes it a document that can be compiled directly using `system` to call a LaTeX compiler.


## Example

### 1. Install and load the package

```{r echo=T, eval=T}
library(textables)
```


### 2. Construct example regression output

```{r echo=T, eval=T}

library(data.table)

dd <- data.table(
    group = c("Full Sample", "Subsample"), 
    coef = c(1.2, 3.42),
    se = c(.6, .481), 
    p = c(.051, .0), 
    N = c(1234567, 891011)
  )

print(dd)

```

### 3. Set up the table columns/alignment

```{r echo=T, eval=T}
alignments <- "lcc"
number_columns <- nchar(alignments)
```


### 4. Make a row with group names using `row_string`

```{r echo=T, eval=T}
group_headers = row_string(" ", dd$group)

print(group_headers)
```



### 5. Three ways to use `row_numeric`, controlling decimals

#### 5.1 Make a row with coefficients and stars using `row_numeric`

```{r echo=T, eval=T}
coefficients = row_numeric("Effect of X on Y", dd$coef, pvalues = dd$p, dec = 3)

print(coefficients)
```


#### 5.2 Make a row with standard errors using `row_numeric`

```{r echo=T, eval=T}
standard_errors = row_numeric(" ", dd$se, dec = 3, se = T)

print(standard_errors)
```


#### 5.3 Make a row with integers using `row_numeric`

```{r echo=T, eval=T}
sample_size = row_numeric("Sample Size (1,000)", dd$N / 1000, dec = 0)

print(sample_size)
```
Note that commas are added automatically.

### 6. Make a header row to span the columns with `header_row`

```{r echo=T, eval=T}
header_across_columns = row_header(" ", "Results at the Worker-level", number_columns - 1)

print(header_across_columns)
```


### 7. Put all of the pieces together, separating with mid-rules


```{r echo=T, eval=T}

tabular <- c(
  header_across_columns,
  group_headers,
  rule_mid_partial(2, number_columns),
  coefficients,
  standard_errors,
  rule_mid(),
  sample_size
)

print(tabular)
```



### 8. Add begin/end with pretty rules


```{r echo=T, eval=T}
tabular = tabular_topbottom_pretty(tabular,alignments)

print(tabular)
```

There is also `tabular_topbottom` that does not add the pretty rules.


### 9. Export as .tex file and compile as example.pdf

Use the `stand_alone` option to wrap the table in begin/end document commands so that the table can be compiled directly by LaTeX.

```{r echo=T, eval=T}
write_tex(tabular,filename='example.tex',stand_alone=T)

system("pdflatex example.tex")
```
See example.pdf, which was compiled by this `system` command.
